---
title: "Collegiate Sports"
output: html_document
---
Our project is based on Collegiate Sports data.

link to the data: https://github.com/rfordatascience/tidytuesday/blob/master/data/2022/2022-03-29/readme.md?fbclid=IwAR1KAJ2uZjaS_Ao7Ocpiz23QlxDJESjrOdzbtH5Ni838aAgviSplkzpdc4A  

We want to test and describe the variance of revenue between two different populations: male and female.  

Our null hypothesis suggests the two populations have equal variances.  

Our alternative hypothesis suggests male revenue has greater variance then female.

$$
H_0: \sigma^2_{\text{Male}} = \sigma^2_{\text{Female}}
$$
$$
H_1: \sigma^2_{\text{Male}} > \sigma^2_{\text{Female}}
$$

```{r warning=FALSE, message=FALSE}
library(tidyverse)
install.packages("tidytuesdayR", repos = "http://cran.us.r-project.org")
tuesdata <- tidytuesdayR::tt_load('2022-03-29')
sports <- tuesdata$sports
```

```{r}
Q1_data <- tibble(value = c(sports$rev_men, sports$rev_women),
                  n.participant = c(sports$sum_partic_men, sports$sum_partic_women), 
                  Gender = rep(c("Men", "Women"), each = nrow(sports))) 
```

```{r}
Q1_data %>% group_by(Gender) %>% summarise(M = mean(value, na.rm = T),
                                           SD = sd(value, na.rm = T),
                                           n = sum(n.participant))

```

Since we need normal distribution for the test residuals, we demonstrated the scale of the variance with logarithmic scale.

```{r}
var.test(x = log1p(sports$rev_women), y = log1p(sports$rev_men), 
         data = sports, ratio = 1, alternative = "less")
```

We reject the null hypothesis with $F_0 = 0.7328$ , and $p-value<0.05$.  

According to the following variance test, our findings accept the alternative hypothesis.  

Male revenue has greater variance then female.

```{r warning=FALSE, message=FALSE}
Q1_data %>% 
  ggplot(aes(x=value, fill = Gender)) + geom_density(alpha = 0.5) + scale_x_log10() +
  labs(x = "Log revenue") + theme_bw()
```

Our next model is Multiple linear regression.  

We want to Describe the relationship between revenue influenced by expenditures interacted with sport category.  

We created 5 main categories and additional one for other sports.  

Baseball is our reference category is.  

```{r message=FALSE}
Q2_data <- tibble(sport = case_when(
  sports$sports == "Baseball"   ~ "Baseball",
  sports$sports == "Soccer"     ~ "Soccer",
  sports$sports == "Football"   ~ "Football",
  sports$sports == "Basketball" ~ "Basketball",
  sports$sports == "Golf" ~ "Golf",
  TRUE                          ~ "Other")) %>% mutate(total_exp_menwomen = sports$total_exp_menwomen,
                                                       total_rev_menwomen = sports$total_rev_menwomen)

Q2_data %>% pivot_longer(-sport) %>% group_by(sport, name) %>% 
  summarise(M = mean(value, na.rm=T),
            SD = sd(value, na.rm=T),
            n = n())
```


```{r}
mod <- lm(total_rev_menwomen ~ total_exp_menwomen * sport, data = Q2_data)
summary(mod)
```
The regression explains great amount of variance as we get high values of $R-square$ and $Adjusted R-square$.  

In addition, the $p-value < 2.2e-16$ which increase the model significance.  

```{r warning=FALSE}
library(scales)
Q2_data %>% ggplot(aes(x=total_exp_menwomen, y=total_rev_menwomen)) +
  geom_point(alpha = 0.3) + 
  geom_smooth(method = "lm", se = F, formula = "y~x") +
  facet_wrap(~sport, scales = "free_x", ncol = 6) + 
  theme_bw() + 
  scale_x_continuous(labels = comma) +
  scale_y_continuous(labels = comma) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.3))+
  labs(x=NULL, y=NULL, title = "Revenue Vs. Exp")
```


As demonstrated,the football expenditures has great influence of the incline.  

The other categories distinguish a linear relationship, but not as pronounced.

We acknowledge Homoscedasticity (the assumption of equal or similar variances) is an important assumption of the regression, although we noticed otherwise. 

